\documentclass[11pt]{article}

% some definitions for the title page
\newcommand{\reporttitle}{example}
\newcommand{\reportdescription}{example description}

% load some definitions and default packages
\input{../../../.latex-templates/includes}
\input{../../../.latex-templates/notation}

\bibliography{../bibliography}

\begin{document}

% Include the title page
\input{../../../.latex-templates/titlepage}

\tableofcontents

\section{Pytorch}

Pytorch will be used for the coursework, recommended is to look at \href{https://pytorch.org/tutorials/beginner/deep_learning_60min_blitz.html}{this} link.

\section{Books}

Recommended NLP Books:

\begin{itemize}
    \item Speech and Language Processing. Dan Jurafsky and James H. Martin~\cite{book-speech-and-language-processing}
    \item A Primer on Neural Network Models for Natural Language Processing.
          Yoav Goldberg~\cite{primer-on-nlp}
    \item Natural Language Processing. Jacob Eisenstein~\cite{git-natural-language-processing}
\end{itemize}

Recommended ML Books:

\begin{itemize}
    \item Artificial Intelligence: a Modern Approach. (2009) Stuart Russell \& Peter Norvig~\cite{AI-modern-approach} with the solutions available at~\cite{AI-modern-approach-slutions}
    \item Machine Learning. (1997) Tom Mitchell~\cite{tom-mitchell-book}
    \item Neural Networks and Deep Learning. Michael A. Nielsen.
    \item Introduction to Deep Learning. Eugene Charniak~\cite{intro-to-dl-eugene-charniak}
    \item Deep Learning by Ian Goodfellow~\cite{Goodfellow-et-al-2016}
\end{itemize}

State of the art NLP:

\begin{itemize}
    \item Papers with code:~\cite{papers-with-code}
    \item  track the progress in Natural Language Processing:~\cite{track-progress-nlp}
\end{itemize}

\printbibliography

\end{document}